Estimating the replicability of psychology experiments after an initial failure to replicate

### 2024-05-02 manuscript/manuscript.Rmd
fix typos, unsuccessful fight with CSL stuff
to render from latex you need to delete the {0}{1} after CSLReferences
... sigh

smd_scale_small <- ggplot(for_plotting_d |> anti_join(big),aes(x=reorder(name_pretty, point, FUN=max),y=point,ymin=low,ymax=high,shape=type, color=type, fill=type, group=desc(type)))+
  geom_errorbar(size=.5,width=0, position=position_dodge(width=.5))+
  geom_point(position=position_dodge(width=.5))+
 coord_flip()+
  scale_size_area()+
  geom_hline(yintercept=0,color="black")+
    scale_color_manual(values=colors, aesthetics=c("color", "fill"))+
  scale_shape_manual(values=shape)+
  theme(legend.position = "bottom", 
        panel.grid.major.x = element_blank(), 
        panel.grid.minor.x=element_blank(),
        legend.title=element_blank(),
        legend.text=element_text(size=10)
        )+
  labs(y="Standarized mean difference", x="")

level_order = c("Yeshurun & Levy 2003",
                "Ngo et al. 2019", 
                 "Krauss & Wang 2003",
                 "Jara-Ettinger et al. 2022")
smd_scale_big <- ggplot(for_plotting_d |> inner_join(big) |> 
                          mutate(name_pretty=factor(name_pretty, levels=level_order)), aes(x=name_pretty,y=point,ymin=low,ymax=high,shape=type, color=type, fill=type, group=desc(type)))+
  geom_errorbar(size=.5,width=0, position=position_dodge(width=.5))+
  geom_point(position=position_dodge(width=.5))+
 coord_flip()+
  scale_size_area()+
  geom_hline(yintercept=0,color="black")+
    scale_color_manual(values=colors, aesthetics=c("color", "fill"))+
    scale_shape_manual(values=shape)+
  theme(legend.position = "none", 
        panel.grid.major.x = element_blank(), 
        panel.grid.minor.x=element_blank()
        )+
  labs(y="Standarized mean difference", x="", )

@@

The rescue of @krauss2003 was successful despite an unsuccessful replication.
This study looked at the influence of guided thinking on whether or not people gave correct justifications (drawn or written) for their answer on the Monty Hall problem.
The original paper reported correct justification from 2/67 (`r round(2/67*100)`%) in the control condition and 13/34 (`r round(13/34*100)`%) in the guided thinking condition.
The first replication struggled to recruit participants who were naive to the problem (an exclusion criterion), and many participants gave very short text responses in the provided text box (only textual responses were allowed).
The replication found 0/8 correct justifications in the control and 0/11 in the guided thinking condition.
While we cannot know for sure what caused the non-replication, there were clear problems observable from the small final sample and low-quality responses.
The rescue targeted these issues by adding a pre-screening that ensured participants were not familiar with the Monty Hall problem, reskinned the problem and changed its name (to reduce googling for answers), and had participants upload drawings for their justifications.
Collectively, we believe these changes brought the rescue closer to the intent of the original.
The rescue had 1/40 (`r round(1/40*100)`%) correct justifications in the control condition and 6/35 (`r round(6/35*100)`%) in the guided thinking condition.
This effect was significant, though smaller than the effect in the original.

@ngo2019 was a second successful rescue. The original study found a large effect; hence the first replication, powering for 80% power on the reported effect, recruited a small sample of 12 people. This small study failed to find the original effect.
The rescue, powered using 2.5x the original sample [as recommended by @simonsohn2015], recovered a clear effect (albeit a much smaller one).
There are reasons to think that some effect sizes in the literature may be inflated [@openscienceconsortium; @camerer2018], and it is also possible that changes to experiments or switches to online could result in noisier samples and thus smaller effect sizes.
Therefore, replications with smaller samples than the original (even if powered to the original effect size), may not be very diagnostic and could potentially benefit from a re-replication.

@@

::: {#refs}
:::

>>>>

smd_scale_small <- ggplot(for_plotting_d |> anti_join(big),aes(x=reorder(name_pretty, point, FUN=max),y=point,ymin=low,ymax=high,shape=type, color=type, fill=type, group=desc(type)))+
  geom_errorbar(size=.5,width=0, position=position_dodge(width=.5))+
  geom_point(position=position_dodge(width=.5))+
 coord_flip()+
  scale_size_area()+
  geom_hline(yintercept=0,color="black")+
    scale_color_manual(values=colors, aesthetics=c("color", "fill"))+
  scale_shape_manual(values=shape)+
  theme(legend.position = "bottom", 
        panel.grid.major.x = element_blank(), 
        panel.grid.minor.x=element_blank(),
        legend.title=element_blank(),
        legend.text=element_text(size=10)
        )+
  labs(y="Standardized mean difference", x="")

level_order = c("Yeshurun & Levy 2003",
                "Ngo et al. 2019", 
                 "Krauss & Wang 2003",
                 "Jara-Ettinger et al. 2022")
smd_scale_big <- ggplot(for_plotting_d |> inner_join(big) |> 
                          mutate(name_pretty=factor(name_pretty, levels=level_order)), aes(x=name_pretty,y=point,ymin=low,ymax=high,shape=type, color=type, fill=type, group=desc(type)))+
  geom_errorbar(size=.5,width=0, position=position_dodge(width=.5))+
  geom_point(position=position_dodge(width=.5))+
 coord_flip()+
  scale_size_area()+
  geom_hline(yintercept=0,color="black")+
    scale_color_manual(values=colors, aesthetics=c("color", "fill"))+
    scale_shape_manual(values=shape)+
  theme(legend.position = "none", 
        panel.grid.major.x = element_blank(), 
        panel.grid.minor.x=element_blank()
        )+
  labs(y="Standardized mean difference", x="", )

@@

The rescue of @krauss2003 was successful despite an unsuccessful replication.
This study looked at the influence of guided thinking on whether or not people gave correct justifications (drawn or written) for their answer on the Monty Hall problem.
The original paper reported correct justification from 2/67 (`r round(2/67*100)`%) in the control condition and 13/34 (`r round(13/34*100)`%) in the guided thinking condition.
The first replication struggled to recruit participants who were naive to the problem (an exclusion criterion), and many participants gave very short text responses in the provided text box (only textual responses were allowed).
The replication found 0/8 correct justifications in the control and 0/11 in the guided thinking condition.
While we cannot know for sure what caused the non-replication, there were clear problems observable from the small final sample and low-quality responses.
The rescue targeted these issues by adding a pre-screening that ensured participants were not familiar with the Monty Hall problem, changed the content items for the problem and changed its name (to reduce googling for answers), and had participants upload drawings for their justifications.
Collectively, we believe these changes brought the rescue closer to the intent of the original.
The rescue had 1/40 (`r round(1/40*100)`%) correct justifications in the control condition and 6/35 (`r round(6/35*100)`%) in the guided thinking condition.
This effect was significant, though smaller than the effect in the original.

@ngo2019 was a second successful rescue. The original study found a large effect; hence the first replication, powering for 80% power on the reported effect, recruited a small sample of 12 people. This small study failed to find the original effect.
The rescue, powered using 2.5x the original sample [as recommended by @simonsohn2015], recovered a clear effect (albeit a much smaller one).
There are reasons to think that some effect sizes in the literature may be inflated [@openscienceconsortium; @camerer2018], and it is also possible that changes to experiments or switches to online could result in noisier samples and thus smaller effect sizes.
Therefore, replications with smaller samples than the original (even if powered to the original effect size), may not be very diagnostic, and could potentially benefit from a re-replication.

@@
